\section{Post-lecture Quiz}
\label{sec:sec:post-quiz}

\subsection{Quiz Questions}

\textbf{Question 1:} Which of the following best describes the relationship between symmetric and asymmetric encryption in a typical secure communication system?

\begin{enumerate}
    \item[A)] Asymmetric encryption is always used for all data transmission
    \item[B)] Symmetric encryption is always used for all data transmission
    \item[C)] \textbf{Asymmetric encryption is used for key exchange, then symmetric encryption is used for data transmission}
    \item[D)] Both types are used simultaneously for the same data
\end{enumerate}

\textbf{Question 2:} In the context of digital certificates and PKI, what is the primary purpose of a Certificate Authority (CA)?

\begin{enumerate}
    \item[A)] To encrypt all network traffic
    \item[B)] \textbf{To verify and digitally sign public key certificates, establishing trust in the public key}
    \item[C)] To store private keys securely
    \item[D)] To generate random encryption keys
\end{enumerate}

\textbf{Question 3:} What is the main security advantage of using SSH key-based authentication compared to password authentication?

\begin{enumerate}
    \item[A)] SSH keys are shorter and easier to remember
    \item[B)] SSH keys provide faster connection speeds
    \item[C)] \textbf{SSH keys eliminate the risk of password interception and provide stronger cryptographic authentication}
    \item[D)] SSH keys work without any network connection
\end{enumerate}

\textbf{Question 4:} In the context of the Trusted Computing Base (TCB), why is it important to minimize the size of the TCB?

\begin{enumerate}
    \item[A)] To reduce system memory usage
    \item[B)] To improve system performance
    \item[C)] \textbf{To reduce the attack surface and limit the number of components that could compromise system security}
    \item[D)] To make the system easier to understand
\end{enumerate}

\textbf{Question 5:} What is the primary difference between integrity and confidentiality in information security?

\begin{enumerate}
    \item[A)] Integrity is about speed, confidentiality is about accuracy
    \item[B)] \textbf{Integrity ensures data has not been modified, while confidentiality ensures data is not disclosed to unauthorized parties}
    \item[C)] Integrity is about authentication, confidentiality is about authorization
    \item[D)] There is no difference between integrity and confidentiality
\end{enumerate}

\subsection{Answer Key}

\begin{enumerate}
    \item \textbf{C} - In practice, asymmetric encryption is used for secure key exchange (due to its key distribution advantages), and then symmetric encryption is used for the actual data transmission (due to its speed advantages).
    
    \item \textbf{B} - Certificate Authorities act as trusted third parties that verify the identity of certificate holders and digitally sign their public key certificates, creating a chain of trust in the PKI system.
    
    \item \textbf{C} - SSH key-based authentication uses cryptographic key pairs instead of passwords, eliminating the risk of password interception and providing much stronger authentication through public-key cryptography.
    
    \item \textbf{C} - The TCB contains all components critical to system security. Minimizing its size reduces the attack surface, meaning fewer components that could potentially be compromised to break system security.
    
    \item \textbf{B} - Integrity ensures data has not been tampered with or corrupted, while confidentiality ensures that sensitive information is not accessed by unauthorized individuals or systems.
\end{enumerate}
